---
title: "DeepDiveR vingette"
author: "Rebecca Cooper"
date: "09/09/2024"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

---
  title: "DeepDiveR vingette"
  output: html_vingette
---
  
  # Introduction
    
  'DeepDiveR' is an R package for estimating biodiversity from fossil occurrence data. It produces formatted input data and a configuration file that can then be exectued in the command line, enabling a clearer and more readily reproducable workflow.
  In this tutorial we will use DeepDiveR to analyse the past diversity of mammalian carnivores using the data of Faurby et al (2019). 
  
# Installation  

To install the library, use the package remotes or by downloading from the GitHub repository:
  
```{r}
# Option 1: install the package from GitHub
devtools::install_github("DeepDive-project/DeepDiveR")
library(DeepDiveR)
```

or...

```{r}
# Option 2: load it from a directory after donwloading it from
# https://github.com/DeepDive-project/DeepDiveR
deepdiver_path <- "path_to_DeepDiveR"
setwd(deepdiver_path)
load_all(".")
library(DeepDiveR)
```

# Preparing data

Load an occurrence table here.
This typically includes one row for each occurrence and 5 columns: taxon name, geographic area, min age, max age, and locality

```{r}
data(carnivora)
dat <- carnivora
```

Next, specify a vector of time bins (they do not need to be equally spaced). Here we use divisions of the stages of the Cenozoic, averaged to 1 My duration:

```{r}
bins <- c(max(dat$MaxAge), 65, 64, 63, 61.6, 60, 59.2, 58.13333, 57.06667, 56, 
          54.975, 53.95, 52.925, 51.9, 50.875, 49.85, 48.825, 47.8, 46.85714, 
          45.91429, 44.97143, 44.02857, 43.08571, 42.14286, 41.2, 40.03667, 
          38.87333, 37.71, 36.7575, 35.805, 34.8525, 33.9, 32.88667, 31.87333, 
          30.86, 29.84667, 28.83333, 27.82, 26.862, 25.904, 24.946, 23.988, 
          23.03, 22.16667, 21.30333, 20.44, 19.3225, 18.205, 17.0875, 15.97, 
          14.895, 13.82, 12.725, 11.63, 10.534, 9.438, 8.342, 7.246, 6.2895, 
          5.333, 4.4665, 3.6, 2.58, 1.8, 0.774, 0.129, 0)
```

Prepare input file for deepdive, set output_file name to save. With the argument 'r=10' we are randomly resampling occurrence ages 10 times from their stratigraphic ranges. For your final analysis, you may want to make more replicates e.g. 100.

```{r}
dd_file_name <- "carnivora_deepdive_input.csv"
prep_dd_input(dat = dat, bins = bins, r = 10, output_file = dd_file_name)
```

# The configuration file

Create a config for the DeepDive analysis. If applicable you can specify the number of living taxa which will be used by the model to calibrate the predicted diversity trajectories.

```{r}
config <- create_config(
  path_wd = path_dat,  
  bins = bins,
  sim_name="carnivora",
  n_areas = length(unique(dat$Area)),
  simulations_file = "simulations_carnivora", 
  models_file = "trained_models_carnivora", 
  present_diversity = 313,
  empirical_input_file = dd_file_name
)
```

To have regions becoming available over time, create area_ages object. Areas are expected to be in alphabetical order.
Here for example we assume that dispersal to South America only becomes possible between 11 and 7 Ma.
Regions disappearing instead of connecting can be made via setting the label argument to: label = "end".

```{r}
area_ages <- rbind(c(max(bins), max(bins)),  # Africa 
                   c(max(bins), max(bins)),  # Asia
                   c(max(bins), max(bins)),  # Europe
                   c(max(bins), max(bins)),  # North America
                   c(11.608, 7.3))           # South America 


areas_matrix(area_ages, n_areas = length(unique(dat$Area)), config, label="start")
```

For the purpose of running a quick test analysis we use low number of training and test simulations.
Note that you will need at least 10,000 training simulations for a real analysis. Entries in the configuration file can be edited in the format below, using the set_value function:

```{r}
set_value(attribute_name = "n_training_simulations", 
          value=100, 
          module="simulations", config)

set_value(attribute_name = "n_test_simulations", 
          value=10, 
          module="simulations", config)
```

Write the configuration file to use in the analysis:

```{r}
config$write("carnivora.ini")
```  

# Executing the analysis

Once the configuration and input files are created, the full DeepDive analysis, inclusive of simulation, model training and empirical predictions, can be carried out through a single command line entered in a Terminal (MacOS and Linux) or Command prompt (Windows) window executing the Python script run_dd_config.py as follows:
  
python run_dd_config.py your_path/config_file.ini

You can additionally specify a working directory where all output files are saved and the number of CPUs used for the parallelized simulations, which will overwrite the corresponding settings in the configuration file, using the flags -wd and -cpu, e.g.

python run_dd_config.py your_path/config_file.ini -wd your_working_directory -cpu 64

This script will create a "simulations" folder containing the training and test sets, and a "trained_models" folder containing the trained models and plots of the training history. This folder will additionally include plots comparing the empirical and simulated fossil features (e.g. number of occurrences through time and per area, number of localities, fraction of singletons, and sampled diversity), CSV files with the predicted diversity trajectories for the test set and for the empirical dataset, and a plot of the estimated diversity trajectory.


# References
Faurby, S., Werdelin, L. & Antonelli, A. (2019) Dispersal ability predicts evolutionary success among mammalian carnivores. BioRxiv, p. 755207.