---
title: "reptilia"
output: html_document
date: "2024-09-26"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r load_packages}
library(DeepDiveR)
library(dplyr)
```

## Prepping Input File

### Loading in Data + EDA

```{r}
data <- read.csv("../BDNN_Arielli/data/reptilia_processed_data/reptilia_deepdive_final.csv", header = TRUE, sep = ",")
head(data)
```

```{r}
sapply(data, class)
```

```{r}
str(data)
```

```{r}
summary(data)
```

Checking if there are nulls

```{r}
data$Area[is.na(data$Area)]
```

```{r}
anyNA(data)
```

We have a lot of unique localities. Let's compare to the default built-in dataset for carnivora in DeepDive

```{r}
length(unique(data[['Locality']]))
```

```{r}
length(unique(data[['Taxon']]))
```

```{r}
length(unique(data[['Area']]))
```

### **Comparing to built-in Carnivora Dataset**

```{r}
data(carnivora)
length(unique(carnivora[['Locality']]))
```

```{r}
length(unique(carnivora[['Taxon']]))
```

```{r}
summary(carnivora)
```

```{r}
head(carnivora)
```

```{r}
length(unique(carnivora[['Area']]))
```

```{r}
summary(carnivora$MaxAge)
summary(carnivora$MinAge)
# According to the Carnivora vignette, the bins start at the max maxage age, and go down to the min minage, so my bins vector should do the same
```

### Defining time bins by stages

```{r}
bins <- c(max(data$MaxAge),290.1, 283.5, 273, 264.3, 259.5, 252, 247, 242, 237, 227, 217)

bins
```

```{r}
range(data$MaxAge)
range(data$MinAge)

```

As you can see, the lowest bin does not include our lowest MinAge. So I added 199.3 to correspond to the Hettangian-Sinemurian border

```{r}
bins <- c(max(data$MaxAge),290.1, 283.5, 273, 264.3, 259.5, 252, 247, 242, 237, 227, 217, 199.3)

bins
```

```{r}
summary(carnivora)
```

The example carnivora dataset has much smaller Age columns because Carnivora is extant. Ran into issues so trouble shooting **below** by shifting my ages and date ranges to present day.

## Trouble Shooting

### Shifting Time Bins/Ages

Ran into issues so trouble shooting **below** by shifting my ages and date ranges to present day. Potentially the large age values are causing issues?

```{r}
data_shifted_ages <- data
data_shifted_ages$MinAge <- data_shifted_ages$MinAge - 199.3
data_shifted_ages$MaxAge <- data_shifted_ages$MaxAge - 199.3

head(data_shifted_ages)

```

```{r}
bins_shifted <- c(max(data_shifted_ages$MaxAge),290.1-199.3, 283.5-199.3, 273-199.3, 264.3-199.3, 259.5-199.3, 252-199.3, 247-199.3, 242-199.3, 237-199.3, 227-199.3, 217-199.3, 0)
bins_shifted
```

```{r}
bins_shifted <- c(99.6, 99, 98, 97, 96, 95, 94, 93, 92, 91, 90,
89, 88, 87, 86, 85, 84, 83, 82, 81, 80,
79, 78, 77, 76, 75, 74, 73, 72, 71, 70,
69, 68, 67, 66, 65, 64, 63, 62, 61, 60,
59, 58, 57, 56, 55, 54, 53, 52, 51, 50,
49, 48, 47, 46, 45, 44, 43, 42, 41, 40,
39, 38, 37, 36, 35, 34, 33, 32, 31, 30,
29, 28, 27, 26, 25, 24, 23, 22, 21, 20,
19, 18, 17, 16, 15, 14, 13, 12, 11, 10,
9, 8, 7, 6, 5, 4, 3, 2, 1, 0)
bins_shifted
```

### Adding a 2nd Area

```{r}
data_shifted_ages[1, 'Area'] <- 2 
head(data_shifted_ages)
```

### Input File

```{r}
setwd("C:/Users/SimoesLabAdmin/Documents/DeepDiveR/reptilia")
# Create input file for DeepDive
prep_dd_input(
  # Specify occurrence data.frame
  dat = data_shifted_ages,
  # Specify vector containing time bin boundaries
  bins = bins_shifted,
  # Specify number of replicates
  r = 100,
  # Specify name of created file
  output_file = "reptilia_deepdive_input_TEST.csv")
```

```{r}
# Create configuration file for DeepDive
config_test <- create_config(
  # Specify name for the simulations
  name = "reptilia_1_TEST",
  # Name of data file
  data_file = "reptilia_deepdive_input_TEST.csv",
  bins = bins_shifted,
  n_areas = length(unique(data_shifted_ages$Area)),
  )
```

```{r}
edit_config(config = config_test,
            module = "general",
            parameter = "present_diversity", 
            value = 0)
```

```{r}
# Set number of training simulations to 100
edit_config(config = config_test,
            module = "simulations",
            parameter = "n_training_simulations", 
            value = 100)

# Set number of test simulations to 10
edit_config(config = config_test,
            module = "simulations",
            parameter = "n_test_simulations", 
            value = 10)

```

```{r}
config_test$write("reptilia_config_1_TEST.ini")

```

The above did not work to solve the error

## Combining into an Input File

```{r create_data, eval = FALSE}
setwd("C:/Users/SimoesLabAdmin/Documents/DeepDiveR/reptilia")
# Create input file for DeepDive
prep_dd_input(
  # Specify occurrence data.frame
  dat = data, # this one does not have a second area
  # Specify vector containing time bin boundaries
  bins = bins_shifted, # these are the un-shifted bins
  # Specify number of replicates
  r = 10,
  # Specify name of created file
  output_file = "reptilia_deepdive_input.csv")

```

## Configuration File

Settings for the analysis/parameters for the simulation

```{r create_config}
setwd("C:/Users/SimoesLabAdmin/Documents/DeepDiveR/reptilia")
# Create configuration file for DeepDive
config <- create_config(
  # Specify name for the simulations
  name = "reptilia_1",
  # Name of data file
  data_file = "reptilia_deepdive_input.csv",
  bins = bins,
  n_areas = length(unique(data$Area)),
  )
```

```{r}
edit_config(config = config,
            module = "general",
            parameter = "present_diversity", 
            value = 0)

# Default # of training sims = 10k, default test sims = 100

```

```{r}
# Set number of training simulations to 100
edit_config(config = config,
            module = "simulations",
            parameter = "n_training_simulations", 
            value = 100)

# Set number of test simulations to 10
edit_config(config = config,
            module = "simulations",
            parameter = "n_test_simulations", 
            value = 10)
```

```{r}
# Write the config file
config$write("reptilia_config_1.ini")
```

## Running DeepDive

We will access the programs in the "deepdive" repo (not this repo) in order to run the simulations. The **Terminal** command:print(config$data$simulations\$carrying_capacity_multiplier)

```{python run, eval = FALSE}
python ../deepdive/run_dd_config.py reptilia/reptilia_config_1.ini -wd reptilia/
```

### Outputs:

-   

-   
