```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r load_packages}
library(remotes)
library(dplyr)
remotes::install_github("DeepDive-project/DeepDiveR")
library(DeepDiveR)

```

## Prepping Input File

### Loading in Data + EDA

```{r}
data <- read.csv("C:/Users/SimoesLabAdmin/Documents/DeepDiveR/updated_occurrences/reptilia_all/rep_all_dd_final.csv", header = TRUE, sep = ",")
head(data)
```

```{r}
# Only keeping select columns
data <- data[, c('Locality', 'Taxon', 'Area', 'MinAge', 'MaxAge')]
```

```{r}
head(data)
```

```{r}
sapply(data, class)
```

```{r}
str(data)
```

```{r}
summary(data)
```

Checking if there are nulls

```{r}
data$Area[is.na(data$Area)]
```

```{r}
anyNA(data)
```

### Defining time bins by stages

Time bins should start at the max MaxAge and go down to just below the min MinAge

```{r}
range(data$MaxAge)
range(data$MinAge)

```


```{r}
bins <- c(max(data$MaxAge),290.1, 283.5, 274.4, 264.3, 259.5, 252, 246.7, 242, 237, 227, 217, 208, 199.3)

bins
```

Ages needed to be shifted to the present, otherwise the simulations would freeze. If using extant taxa, this is not necessary obviously

```{r}
data_shifted_ages <- data
data_shifted_ages$MinAge <- data_shifted_ages$MinAge - 199
data_shifted_ages$MaxAge <- data_shifted_ages$MaxAge - 199

head(data_shifted_ages)

```

```{r}

# Actually, just made last bin = 0 to mimic present day. 199.3 - 199 would've been basically 0 anyway
bins_shifted <- c(max(data_shifted_ages$MaxAge),290.1 - 199, 283.5 - 199, 274.4 - 199, 264.3 - 199, 259.5 - 199, 252 - 199, 246.7 - 199, 242 - 199, 237 - 199, 227 - 199, 217 - 199, 208 - 199, 0)
bins_shifted

```
## There's an error in the program. Instructions say to use 'Area' column, but code requires it to be called 'Region'. So, renaming 'Area' to 'Region' in the data frame before calling prep_dd_input

```{r}
# rename 'Area' to 'Region'
data_shifted_ages <- data_shifted_ages %>%
  rename(Region = Area)
```


### Calling prep_dd_input

```{r}
setwd("C:/Users/SimoesLabAdmin/Documents/DeepDiveR/updated_occurrences/reptilia_all/")
# Create input file for DeepDive
prep_dd_input(
  # Specify occurrence data.frame
  dat = data_shifted_ages,
  # Specify vector containing time bin boundaries
  bins = bins_shifted,
  # Specify number of replicates
  r = 100,
  # Specify name of created file
  output_file = "rep_all_deepdive_input.csv")
```

```{r}
# Create configuration file for DeepDive
config_test <- create_config(
  # Specify name for the simulations
  name = "rep_all",
  # Name of data file
  data_file = "rep_all_deepdive_input.csv",
  bins = bins_shifted,
  n_regions = length(unique(data_shifted_ages$Region)),
  )
```

```{r}
# Set number of training simulations to 100
edit_config(config = config_test,
            module = "simulations",
            parameter = "n_training_simulations", 
            value = 10000)

# Set number of test simulations to 10
edit_config(config = config_test,
            module = "simulations",
            parameter = "n_test_simulations", 
            value = 1000)

edit_config(config = config_test,
            module = "simulations",
            parameter = "extant_sp", 
            value = 0)

edit_config(config = config_test,
            module = "simulations",
            parameter = "pr_extant_clade", 
            value = 0)

# Optional: p_mass_extinction parameter sets the probability of a mass extinction per my
# My study spans ~100 mys, so a p_mass_extinction = 0.02/my = 2 mass extinctions in the
# time frame of my study, which is about what we're expecting
# Default value is 0.01
edit_config(config = config_test,
            module = "simulations",
            parameter = "p_mass_extinction", 
            value = 0.02)


```

```{r}
config_test$write("rep_all_config.ini")

```

## Calling Config: Terminal Command

We will access the programs in the "deepdive" repo (not this repo) in order to run the simulations. The **Terminal** command:print(config$data$simulations\$carrying_capacity_multiplier)

```{python run, eval = FALSE}
python ../deepdive/run_dd_config.py reptilia/reptilia_config.ini -wd reptilia
```

### Outputs:

-   rep

-   
