---
title: "synapsida"
output: html_document
date: "2024-09-26"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r load_packages}
library(remotes)
library(DeepDiveR)
library(dplyr)

remotes::install_github("DeepDive-project/DeepDiveR", ref = "application_note")
```

## Prepping Input File

### Loading in Data + EDA

```{r}
data <- read.csv("../../BDNN_Arielli/data/synapsida_processed_data/synapsida_deepdive_final.csv", header = TRUE, sep = ",")
head(data)
```

```{r}
# Only keeping select columns
data <- data[, c('Locality', 'Taxon', 'Area', 'MinAge', 'MaxAge')]
```

```{r}
head(data)
```

```{r}
sapply(data, class)
```

```{r}
str(data)
```

```{r}
summary(data)
```

Checking if there are nulls

```{r}
data$Area[is.na(data$Area)]
```

```{r}
anyNA(data)
```

**Comparing to default 'carnivora' dataset used in vignettes**

```{r}
length(unique(data[['Locality']]))
```

```{r}
data(carnivora)
length(unique(carnivora[['Locality']]))
```

```{r}
length(unique(data[['Taxon']]))
```

```{r}
length(unique(carnivora[['Taxon']]))
```

```{r}
length(unique(data[['Area']]))
```

```{r}
head(carnivora)
```

```{r}
length(unique(carnivora[['Area']]))
```

```{r}
table(carnivora$Area)

```

```{r}
summary(carnivora$MaxAge)
summary(carnivora$MinAge)
# According to the Carnivora vignette, the bins start at the max maxage age, and go down to just below the minimum minage, so my bins vector should do the same
```

### Defining time bins by stages

```{r}
bins <- c(max(data$MaxAge),290.1, 283.5, 273, 264.3, 259.5, 252, 247, 242, 237, 227, 217, 208)

bins
```

```{r}
range(data$MaxAge)
range(data$MinAge)

```

As you can see, the lowest bin does not include our lowest MinAge. So I added 199.3 to correspond to the Hettangian-Sinemurian border

```{r}
bins <- c(max(data$MaxAge),290.1, 283.5, 273, 264.3, 259.5, 252, 247, 242, 237, 227, 217, 208, 199.3)

bins
```

```{r}
summary(carnivora)
```

The example carnivora dataset has much smaller Age columns because Carnivora is extant

During my test runs of reptilia (in test/reptilia), I found ages needed to be shifted to the present, otherwise the simulations would freeze

### Shifting Time Bins/Ages

```{r}
data_shifted_ages <- data
data_shifted_ages$MinAge <- data_shifted_ages$MinAge - 199
data_shifted_ages$MaxAge <- data_shifted_ages$MaxAge - 199

head(data_shifted_ages)

```

```{r}
bins_shifted <- c(max(data_shifted_ages$MaxAge),290.1-199, 283.5-199, 273-199, 264.3-199, 259.5-199, 252-199, 247-199, 242-199, 237-199, 227-199, 217-199, 208-199, 0)
# Actually, just made last bin = 0 to mimic present day. 199.3 - 199 would've been basically 0 anyway

bins_shifted
```

### Calling prep_dd_input

```{r}
setwd("C:/Users/SimoesLabAdmin/Documents/DeepDiveR/synapsida/")
# Create input file for DeepDive
prep_dd_input(
  # Specify occurrence data.frame
  dat = data_shifted_ages,
  # Specify vector containing time bin boundaries
  bins = bins_shifted,
  # Specify number of replicates
  r = 100,
  # Specify name of created file
  output_file = "synapsida_deepdive_input.csv")
```

```{r}
# Create configuration file for DeepDive
config_test <- create_config(
  # Specify name for the simulations
  name = "synapsida",
  # Name of data file
  data_file = "synapsida_deepdive_input.csv",
  bins = bins_shifted,
  n_areas = length(unique(data_shifted_ages$Area)),
  )
```

```{r}
edit_config(config = config_test,
            module = "general",
            parameter = "present_diversity", 
            value = 0)
```

```{r}
# Set number of training simulations to 100
edit_config(config = config_test,
            module = "simulations",
            parameter = "n_training_simulations", 
            value = 10000)

# Set number of test simulations to 10
edit_config(config = config_test,
            module = "simulations",
            parameter = "n_test_simulations", 
            value = 1000)

```

```{r}
config_test$write("synapsida_config.ini")

```

## Calling Config: Terminal Command

We will access the programs in the "deepdive" repo (not this repo) in order to run the simulations. The **Terminal** command:print(config$data$simulations\$carrying_capacity_multiplier)

```{python run, eval = FALSE}
python ../deepdive/run_dd_config.py synapsida/synapsida_config.ini 
```

### Outputs:

-   rep

-   
